Social inequality in mortality among adults and elderly in northern Sweden 1851--2013\footnote{Paper presented at the session 'Health and Mortality in Arctic Populations'. The Ninth International Congress on Arctic Social Sciences, Umeå, June 8--12, 2017}

### 2017-05-19 icass.Rnw

\author{Göran Broström\footnote{email: \texttt{goran.brostrom@umu.se}} and
    Sören Edvinsson\footnote{email: \texttt{soren.edvinsson@umu.se}} \\
Centre for Demographic and Ageing Research ({\sc Cedar})\\
Umeå University, SE-90187 Umeå, Sweden}

\title{Social inequality in mortality among
  adults and elderly in northern Sweden 1851--2013\footnote{Paper presented
    at the
    session 'Socoeconomic inequalities and mortality I: Europe'. 41st Annual Meeting of the Social Science History Association, 
    Chicago, Illinois
    November 17--20, 2016}}

\date{November 19, 2016}

@@

<<readtab>>=
load("data/vbTab.rda") 
vbTab$period <- factor(vbTab$period)
levels(vbTab$hisclass40) <- c("elite", "lowMan", "lowWhiteC", "worker")
levels(vbTab$hisclass60) <- c("elite", "lowMan", "lowWhiteC", "worker")
levels(vbTab$period)[8] <- "1991-2001"
load("data/vb.rda")
vb$period <- factor(vb$period)
levels(vb$hisclass) <- c("elite", "lowMan", "lowWhiteC", "worker")
levels(vb$hisclass40) <- c("elite", "lowMan", "lowWhiteC", "worker")
levels(vb$hisclass60) <- c("elite", "lowMan", "lowWhiteC", "worker")
levels(vb$period)[8] <- "1991-2001"
@

@@

<<nana, results = 'asis'>>=
library(xtable)
load("data/vb0.rda")
vb0$hisclass <- factor(vb0$hisclass, labels = c("elite", "lowMan", "lowWhiteC", "worker"))
x <- with(vb0, table(period, hisclass, useNA = "ifany"))
x <- round(prop.table(x, 1) * 100)
xtable(x, caption = "Hisclass by time period, per cent.", label = "tab:nana", digits = 0)
@

>>>>

\author{Sören Edvinsson\footnote{email: \texttt{soren.edvinsson@umu.se}} and
Göran Broström\footnote{email: \texttt{goran.brostrom@umu.se}}
     \\
Centre for Demographic and Ageing Research ({\sc Cedar})\\
Umeå University, SE-90187 Umeå, Sweden}

\title{Social inequality in mortality among
  adults and elderly in northern Sweden 1851--2013\footnote{Paper presented
    at the
    session 'Health and Mortality in Arctic Populations'. The Ninth International
    Congress on Arctic Social Sciences,
    Umeå,
    June 8--12, 2017}}

\date{May 18, 2017}

@@

<<readtab>>=
load("data/vbTab.rda") 
vbTab$period <- factor(vbTab$period)
levels(vbTab$hisclass40) <- c("elite", "lowMan", "lowWhiteC", "worker")
levels(vbTab$hisclass60) <- c("elite", "lowMan", "lowWhiteC", "worker")
levels(vbTab$period)[8] <- "1991-2001"
load("data/vb0.rda")
vb0$period <- factor(vb0$period)
vb0$hisclass[is.na(vb0$hisclass)] <- 4
vb0$hisclass <- factor(vb0$hisclass, labels = c("elite", "middle", "worker", "NA"))
vb <- vb0
@

@@

<<nana, results = 'asis'>>=
library(xtable)
##load("data/vb0.rda")
##vb0$hisclass <- factor(vb0$hisclass, labels = c("elite", "lowMan", "lowWhiteC", "worker"))
x <- with(vb0, table(period, hisclass, useNA = "ifany"))
x <- round(prop.table(x, 1) * 100)
xtable(x, caption = "Hisclass by time period, per cent.", label = "tab:nana", digits = 0)
@

### 2017-05-19 icass2.Rnw

<<readtab>>=
load("data/vbTab.rda") 
vbTab$period <- factor(vbTab$period)
levels(vbTab$hisclass40) <- c("elite", "lowMan", "lowWhiteC", "worker")
levels(vbTab$hisclass60) <- c("elite", "lowMan", "lowWhiteC", "worker")
levels(vbTab$period)[8] <- "1991-2001"
load("data/vb0.rda")
vb0$period <- factor(vb0$period)
vb0$hisclass[is.na(vb0$hisclass)] <- 4
vb0$hisclass <- factor(vb0$hisclass, labels = c("elite", "middle", "worker", "NA"))
vb <- vb0
@

@@

<<nana, results = 'asis'>>=
library(xtable)
##load("data/vb0.rda")
##vb0$hisclass <- factor(vb0$hisclass, labels = c("elite", "lowMan", "lowWhiteC", "worker"))
x <- with(vb0, table(period, hisclass, useNA = "ifany"))
x <- round(prop.table(x, 1) * 100)
xtable(x, caption = "Hisclass by time period, per cent.", label = "tab:nana", digits = 0)
@

>>>>

<<readtab>>=
load("data/vb0.rda")
vb0$hisclass[is.na(vb0$hisclass)] <- 4
vb0$hisclass <- factor(vb0$hisclass, labels = c("elite", "middle", "worker", "NA"))
plot(c(1851, 1851, 1951, 1951, 1851), c(40, 90, 90, 40, 40),
     type = "l", xlim = c(1851, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(0, 90))
axis(1, at = c(1851, 1901, 1951, 1961, 1981, 2014), las = 2)
axis(2, las = 2, at = c(0, 40, 65, 90))
axis(4, las = 2, at = c(0, 40, 65, 90))
box()
abline(h = 0)
lines(c(1961, 1961, 2014, 2014, 1961), c(40, 90, 90, 40, 40),
      col = "blue", lwd = 2)
#
n.rec <- NROW(vb0[vb0$dataset == "poplink", ] )
n.pl <- with(vb0[vb0$dataset == "poplink", ], length(unique(id)))
n.d <-  with(vb0[vb0$dataset == "poplink", ], sum(event))
text(1900, 70, "PopLink", cex = 2, col = "red")
text(1899, 60, paste(n.rec, "records"))
text(1900, 55, paste(n.pl, "persons"))
text(1899.5, 50, paste(n.d, "deaths"))
#
n.rec <- NROW(vb0[vb0$dataset != "poplink", ] )
n.pl <- with(vb0[vb0$dataset != "poplink", ], length(unique(id)))
n.d <-  with(vb0[vb0$dataset != "poplink", ], sum(event))

text(1985, 70, "Linnaeus", cex = 2, col = "blue")
text(1984, 60, paste(n.rec, "records"))
text(1985.2, 55, paste(n.pl, "persons"))
text(1985.2, 50, paste(n.d, "deaths"))
@

@@

<<nana, results = 'asis'>>=
library(xtable)
load("data/vb0.rda")
vb0$hisclass <- factor(vb0$hisclass, labels = c("elite", "lowMan", "lowWhiteC", "worker"))
x <- with(vb0, table(period, hisclass, useNA = "ifany"))
x <- round(prop.table(x, 1) * 100)
xtable(x, caption = "Hisclass by time period, per cent.", label = "tab:nana", digits = 0)
@
